Allee Effect and Cultural Evolution

### 2015-06-05 manuscript/draft.tex
updated plots

\begin{abstract}
Several social learning strategies are guided by an evaluation of alternative cultural variants by means of the expected payoffs associated by these. The assumption of the social learner is that the payoff received by the transmitter while expressing a given cultural trait, is a proxy for for inferring the qualitative features of the trait itself. Here we explore the theoretical implications when payoff directly affects reproductive fitness and it is density dependent, that is its expected amount is a function of the number of individuals possessing the same trait. More specifically we consider situations where the density dependence is simultaneously direct and inverse, whereby the highest payoff is achieved at intermediate densities.This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits that are enhanced by cooperation or mutual facilitation, but are somehow constrained by some carrying capacity.  We  explores the evolutionary dynamics of a single trait model with two variants, each exhibiting a different Allee effect. We used a modified Lotka-Volterra model of payoff-biased social learning, and show that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion, stable co-existence, and fixation to single variant.

>>>>

\begin{abstract}
Several social learning strategies are guided by an evaluation of alternative cultural variants by means of the expected payoffs associated by these. The assumption of the social learner is that the payoff received by the transmitter while expressing a given cultural trait, is a proxy for for inferring the qualitative features of the trait itself. Here we explore the theoretical implications when payoff directly affects reproductive fitness and it is density dependent, that is its expected amount is a function of the number of individuals possessing the same trait. More specifically we consider situations where the density dependence is simultaneously direct and inverse, whereby the highest payoff is achieved at intermediate densities. This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits that are enhanced by cooperation or mutual facilitation, but are somehow constrained by some carrying capacity.  We  explores the evolutionary dynamics of a single trait model with two variants, each exhibiting a different Allee effect. We used a modified Lotka-Volterra model of payoff-biased social learning, and show that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion, stable co-existence, and fixation to single variant.
